A Handbook for Tenured Track Faculty in Computer Science in the US

### 2024-07-11 faculty.tex
updates

\section{Types of Funding}
\subsection{NSF and other Government Funding}

\paragraph{Other Government Funding}

@@

\subsection{Industrial Gifts}

@@

\chapter{Mentoring}
\section{Mentoring Undergraduate Students}
\section{Mentoring Graduate Students}
\section{Mentoring Postdocs}
\section{Mentoring Junior Faculty}

@@

Officially services are only counted for 10\% of your work, however they are very important for your tenure and career.
%You are expected to have certain types of services as a faculty member.  
Some services are required (e.g., when your department chair assigns you to some committees), but some are not and can be declined (e.g., an invitation to be a PC member).

@@

Toward the end of your 3rd year (or early 4th year), you will need to submit a ``midterm'' \emph{renewal} application to renew your contract for the 4th year (and perhaps also 5th year).
While somewhat similar to tenure evaluation, this is less intense and  does not require reference letters or presentation.
%The department will then review your application and vote on whether to renew you for your 4th (and perhaps also 5th year). 
Your mentor presents and argues for your case in the dept. meeting, but the decision is made within the dept. and usually does not go up to the college or university.  It is still signed by the Dean, who often just follows the dept's recommendation.

While the renewal is straightforward for most people, it can be surprise and shock to some.

>>>>

\section{Types of Funding}
\paragraph{NSF} NSF is the obvious funding source. Among many good things about NSF, two stand out: (transparency) proposals are evaluated by an independent panel and are based on merit rather than personal connections; (autonomy) the flexibility and freedom given to focus on pure research without the pressure of producing concrete deliverables.  Because of these reasons, NSF funding is often considered the prestigious and competitive, even for very experienced senior researchers from top places (which don't mean anything). Note that a PI can use at most 2 months of summer salary from NSF grants (if you need the 3rd summer month, you need to find other sources of funding).

\paragraph{DARPA, ONR, and other defense sponsors} DARPA, ONR, Army, Airforce, etc are like a contract, with deliverables required at regular intervals. Contracts are awarded in phases to multiple teams, and funding is cut for underperforming teams. This puts a lot of pressure on both the PIs and their students, requiring them to spend significant time on reports, meetings, and product development (e.g., creating hacker-proof software rather than research prototypes). Additionally, in many cases funding often depends on existing connections or being affiliated with institutions that frequently receive such awards.

However, these sponsors are good sources of funding and can provide the additional resources (e.g., summer months for faculty). Note that some defense programs, e.g., Young Investigator Award, are more like NSF grants, with less stringent requirements and no deliverables required.

@@

\subsection{Industrial Gifts}
As its name suggests, industrial gifts are unrestricted research funding from companies such as Amazon, Microsoft, Facebook, and Google. These are highly competitive and prestigious (e.g., the solicitation is open worldwide, and you compete with top researchers from top places).  They are the best kind of funding for three reasons: (1) indirect costs or overheads are typically not levied on these gifts (so you get most of the money), (2) there is no associated timeline by when you should spend them (they give it to you and forget about it), and (3) you can use them for a broader range of things than university and grant/contract funding (e.g., fun events for your research group).

@@

\chapter{Mentoring}

Successfully mentoring students is a key and most rewarding part of your job as a faculty member. 
There will be times when you feel discouraged and frustrated, but there will also be times when you feel proud and happy.  
You will be expected to mentor undergraduate students and graduate students. You will likely need to graduate at least one PhD student to get tenure.

\section{Mentoring Graduate Students}

\section{Mentoring Undergraduate Students}
Surprisingly, some of the best research projects I have seen come from undergraduate students.

I have worked with many undergrads (15+ since 2016). My success rate (publish a research paper with me) with undergrads

\section{Mentoring Postdocs}
\section{Mentoring Junior Faculty}

@@

Officially services are counted for only 10\% of your work. However, they are very important for your tenure and career.
%You are expected to have certain types of services as a faculty member.  
Some services are required (e.g., when your department chair assigns you to some committees), but some are not and can be declined (e.g., an invitation to be a PC member).

@@

Toward the end of your 3rd year (or early 4th year), you will need to submit a ``midterm'' \emph{renewal} application to renew your contract for the 4th year (and perhaps also 5th year).
While somewhat similar to tenure evaluation, this is less intense and  does not require reference letters or presentation.
%The department will then review your application and vote on whether to renew you for your 4th (and perhaps also 5th year). 
Your mentor will also present and argue for your case in the dept. meeting, but the decision is made within the dept. and usually does not go up to the college or university.  It is still signed by the Dean, who often just follows the dept's recommendation.

While the renewal is straightforward for most people, it can be surprise and shock to some. Ideally, it should not be a surprise, as your annual evaluation should give you some idea on how you are doing, e.g., if you are not doing well, the annual evaluation should tell you that you need to improve. However, it can be a surprise due to various reasons such as new dept. chair or administrative changes who have different expectations.
